Boson-Sampling with Single Photons: A Roadmap to Post-Classical Computation

### 2016-09-22 paper.tex
completed proof read, waiting on app.

%\begin{figure}[!htb]
%\includegraphics[width=\columnwidth]{lossy}
%\caption{???} \label{fig:lossy}
%\end{figure}

@@

\begin{abstract}
Boson-sampling is thus far the most feasible approach for demonstrating the supremacy of quantum computation over classical computation in the near future. We take the view that proof-of-principle demonstration of quantum supremacy is a pressing goal, with enormous implications, and should be considered an utmost priority by the experimental community. Here we strive to outline a roadmap to achieving this elusive goal, examining competing architectures and technologies, and paving the way forward for the first experimental demonstration of post-classical computation. We firstly develop a practical model to compare the performance of boson-samplers and classical simulators. In this model, we find that low-photon-number boson-sampling operating in the current experimental regime (\mbox{$n<10$} photons) could beat present commodity computers with GHz clock-rates only if the overall system efficiency and repetition rate are sufficiently high. Further, we calculate the required efficiency and photon-number to surpass classical computers of different clock-rates using various experimental schemes. Especially, we find that only if both photon-number-resolving detectors and high-speed optical shutters are available, scattershot boson-sampling is feasibly scalable to tens of photons. In order to provide a clear roadmap for boson-sampling, we also review and discuss various state-of-the-art single-photon sources, quantum circuits, and single-photon detectors.
\end{abstract}

@@

Quantum computers are expected to be able to solve certain problems much faster than classical computers \cite{bib:1}, and people have been making unremitting efforts towards them for more than two decades. However, actual quantum computers are presently still in their infancy. Thanks to its simplicity, boson-sampling is expected to be the first specialised quantum computer outperforming classical ones in the near term \cite{bib:2}. More importantly, this would also constitute the strongest evidence yet against a foundational tenet in computer science: the Extended Church-Turing Thesis (ECT), which postulates that all realistic physical systems can be efficiently simulated with a (classical) probabilistic Turing machine. However, the realisation of large-scale boson-samplers still faces many demanding technological challenges, despite being far easier to build than universal quantum computers.

To scale up to a large enough level to exhibit quantum supremacy, the first and most important consideration is the total efficiency of the system, including the generation efficiency of single-photon sources, the coupling and propagation efficiency of the circuit, and the coupling and detection efficiency of single-photon detectors.

In Sec.~\ref{sec:sampling_time}, taking the repetition rate of the sampling system into account, we develop a rigorous and fair model for comparing the performance of boson-samplers and classical simulators. Given a certain classical computer, the regime of quantum supremacy of boson-sampling is specified by a simple continuous curve in the plane of total efficiencies versus photon-numbers. \comment{To our surprise, there is such a (optimal) turning point in this curve where four-photon boson-samplers with repetition rates of 76MHz could beat present commodity computers with GHz clock-rates}. Based upon this, such a computer could only be defeated by \mbox{$20\sim 30$}-photon boson-sampling \cite{bib:2, bib:3}. Furthermore, in Sec.~\ref{sec:exp_schemes}, we calculate the required efficiencies and photon-numbers for different experimental schemes to outperform classical computers of given clock-rates. In particular, we find that both photon-number-resolving detectors and high-speed optical shutters are indispensable to correctly implementing scattershot boson-sampling with tens of photons.

@@

\section{Sampling Time} \label{sec:sampling_time}

@@

\subsection{Quantum Simulation}

Boson-sampling is a process in which $n$ indistinguishable photons are injected into an $m$-mode linear optical interferometer, characterised by linear optics transformation $U$ (i.e beamsplitters and phase-shifters), and then detected at output ports using photodetectors, as shown in Fig.~\ref{fig:model} \cite{bib:2}. $U$ evolves the photonic creation operators according to,
\begin{align}
\hat{U} \hat{a}^\dag_i \hat{U}^\dag \to \sum_{j=1}^m U_{i,j} \hat{a}^\dag_j,
\end{align}
where $\hat{a}^\dag_i$ is the photon creation operator for the $i$th mode.

Choosing the mode occupation-number configurations as the basis states, the input state can be written as a list of occupation numbers \mbox{$\ket{S}=\ket{s_1,\dots,s_m}$}, where $s_i$ is the number of photons in the $i$th input mode. As photon-number is preserved via linear optics transformations, the output state can be denoted in the same basis, but a superposition of all different output configurations,
\begin{align}
\ket\varphi = \sum_{i=1}^{|\Phi|}\alpha_i \ket{T_i},
\end{align}
where \mbox{$\ket{T_i} = \ket{t_1,\dots,t_m}$}, and with each basis state satisfying
\begin{align}
\sum_{i=1}^m s_i = \sum_{i=1}^m t_i = n \,\,\forall\,\, S,T.
\end{align}
We define $\Phi$ as the set of all photon-number configuration basis vectors. It is not hard to prove that the total number of elements in $\Phi$ is,
\begin{align}
|\Phi|=\binom{m+n-1}{n}.
\end{align}
The amplitude associated with each output basis is given by \cite{bib:2, bib:12},
\begin{align}
\bra{T_i} \Phi(\hat{U}) \ket{S} = \frac{\mathrm{Per}(U_{S,T_i})}{\sqrt{s_1!\dots s_m!t_1!\dots t_m!}},
\end{align}
where $\Phi(U)$ is a homomorphism of $U$, and $U_{S,T_i}$ is an $n\times n$ submatrix of $U$, obtained by taking $s_i$ copies of the $i$th row of $U$, and $t_i$ copies of the $i$th column of $U$. It has been proven that \cite{bib:2, bib:13, bib:14} one will observe a collision (i.e. two or more photons in the same mode) with probability bounded away from 0 only if the number of modes scales as at least $m=O(n^2)$ -- the so-called `bosonic birthday paradox' \cite{???}. In this instance, it suffices to use non-photon-number-resolving detectors, substantially simplifying technological requirements. Thus, throughout this paper we only discuss the `collision-free' situation, i.e. $s_i,t_j=\{0,1\}$.

@@

\subsection{Classical Simulation}

How can a classical computer simulate this bosonic process, and sample from its exact or approximate probability distribution? A na{\"i}ve brute-force method is to calculate all $|\Phi|$ permanents of the set of submatrices $U_{S,T_i}$, thereby determining the full probability distribution, and then sample from this known distribution. \textbf{However, neither does the boson-sampler know even the approximate probability distribution before enough samples are obtained} \comment{What does that sentence mean?}. Smarter classical algorithms exist to generate correct samples without knowing the entire underlying distribution. For example, using the acceptance-rejection method \cite{bib:15} one can sample an output configuration $\ket{T_i}$ uniformly from the configuration space and a number $u$ from the uniform distribution over $\{0,1\}$, then calculate the permanents of the respective submatrices and compare them to $u$; If \mbox{$u<|\mathrm{Per}(U_{S,T_i})|^2$}, we accept the sample $\ket{T_i}$; If not, it is rejected, and we repeat. Such methods could be successful with probability $O(1)$. One sample only requires calculating one permanent, significantly better than the brute-force method of calculating the entire distribution.

The permanent is a matrix function originating from the permutation symmetry of bosons, similar to the role of determinants for fermions. Na{\"i}ve calculation of the permanent of an $n\times n$ matrix requires $n!$ operations. However, a more efficient classical algorithm developed by Ryser \cite{bib:16} in 1963 requires $O(n^2 2^{n+1})$ operations, which is nonetheless exponential. No general polynomial-time algorithms have been described, in contrast to the clever Gaussian elimination method for calculate matrix determinants. In fact, calculating the permanent was proved to be \#P-complete in 1979 by Valiant \cite{bib:17}. A more intuitive understanding is that the permanent corresponds to counting the number of different matching ways between input state and output state through the interferometer, much like adding Feynman paths \cite{???} to determine quantum amplitudes.

@@

Comparing the requisite times, we plot the critical photon-number as a function of efficiency $\eta$ for the quantum boson-sampler to beat its classical counterpart, as shown in Fig.~\ref{fig:curves_1}(a). The right part of each line is the area of quantum supremacy. For a given single-photon efficiency, with increasing $n$, the quantum simulator's superiority becomes more prominent. Thus, scaling up the size of boson-sampling is essential to achieving quantum supremacy. But there is an upper limit. Taking \mbox{$\eta=0.4$} as an example, it is impossible to reach quantum supremacy for a classical computer of petaFLOPS by only enlarging the number of single photons. A more convenient approach is to increase the efficiency of the quantum system. As the efficiency increases, the minimum photon-number required to achieve quantum supremacy decreases, simplifying experimental resource requirements.

@@

The repetition rate $R$ is another important factor affecting the performance of quantum simulators that we must consider in practical experiments. If $\eta^n$ is taken as the post-selection success probability of a single trial, then \mbox{$\eta^n R$} indicates how many successful runs we can implement per second. For example, to beat a gigaFLOPS classical computer, four-photon boson-sampling with $R=76\mathrm{MHz}$ ($R$ vs $A$) is sufficient if the total efficiency is \mbox{$\eta=0.4$}. As shown in Fig.~\ref{fig:curves_1}(b), however, \mbox{$\eta=0.2$} and \mbox{$n=2$} is sufficient if the repetition rate increases to 1GHz. In other words, by increasing the repetition rate, the efficiency and photon-number requirements are relaxed. However, we cannot increase the repetition rate indefinitely. It is restricted by the pulse rate of the SPS, experimental architecture, and dead-time of the detectors. In the next section we discuss the impact of the choice of experimental architecture.

Here we must point out two things. First, Ryser's algorithm is actually not faster than using the original formula to calculate the matrix permanent when photon number is small (\mbox{$n<6\sim 7$}). Thus the above conditions may be not tight. For example, the efficiency must be larger than $\sim 0.6$ if two- (four)-photon boson-sampling with a repetition rate of 1GHz (76MHz) is expected to beat a gigaFLOPS classical computer.

Why then should we nonetheless scale up to more photons (\mbox{$n=20\sim 30$}), even though boson-sampling with relatively few photons could beat current prevailing personal computers in realistic experimental regimes? One reason is that no finite experiment can make a decisive conclusion about the ECT, since it is a conjecture about the asymptotic limit. In the field of computational complexity, the significance of boson-sampling rests largely on its exponentially faster sampling speed than that of classical simulators, rather than certain times of specific triumphs in the competitions between quantum and classical computers. Thus, through upscaling we can make stronger statements about asymptotic behaviour (albeit never a completely conclusive one). Another cause is that people have not physically constructed relatively large multi-photon interferometers yet. The theory of boson-sampling still faces the risk of failure and quantum devices may break down with more photons. Finally, although no practical uses for boson-sampling have been described, it is entirely plausible that useful applications may be found in the future. Were that the case, scaling up boson-sampling experiments would similarly scale up the size of the problems it could solve.

\section{Experimental Schemes} \label{sec:exp_schemes}

\subsection{Spatial Encoding Architecture}

A fully spatially-encoded scheme is one whereby the circuit itself does not consume any time resources (beyond the propagation time from the input to the output time of the circuit, which is fixed). This includes optical waveguides, fused fibre couplers and bulk-optics circuits, as shown in Fig.~\ref{fig:space_encoding}. The time required for this scheme is the same as Eq.~\ref{eq:tau_q}.

@@

\subsection{Time-Bin Encoding Architecture}

@@

\begin{figure}[!htb]
\includegraphics[width=\columnwidth]{loop_arch}
\caption{The fibre-loop scheme for boson-sampling. A pulse-train of photons in $m$ time-bins defines the input state, with time-bin separation $\tau$. The pulse-train is coupled into and out of the outer loop using two outer dynamic switches. The inner loop of length $\tau$ allows neighbouring time-bins to interfere at the central dynamic switch. The outer two switches need only switch between being completely reflective and completely transmissive, whereas the inner switch must be able to tune to any beamsplitter reflectivity. The only component that scales with the size of the interferometer is the length of the outer loop, which must be long enough to contain the entire pulse-train, $>m\tau$.} \label{fig:loop_arch}
\end{figure}

In this design, a sequence of single photons is injected and interference with one other through the time delay and fibre beamsplitter. As in Fig.~\ref{fig:loop_arch}, a small fibre loop of round-trip time \mbox{$\tau=1/R$} acts as the time delay, where the node is a dynamic beam splitter for interference. The small loop is embedded in a large fibre loop of travelling time $>m\tau$, which is used to return the `photon pulse-train' back for next round of interference. These two embedding fibre loops together fully construct the interference network that encodes only in the temporal degree of freedom. According to Reck's result, one needs $O(m^2)$ beamsplitters to construct an arbitrary \mbox{$m\times m$} interferometer. For a with $m$ time bins, the beamsplitter works $m$ times in each round trip. So $m$ round trips are required to implement an arbitrary  unitary transformation. After $m$ loops of evolution, the photons are all coupled out of the loop for time-resolved coincidence measurements.

As described above, the circuit itself consumes time resources $m^2\tau$, thus the effective repetition rate is,
\begin{align}
\frac{1}{m^2\tau}=\frac{R'}{m^2},
\end{align}
where $R'$ is the repetition frequency of the photonic time-bins. We use it instead of $R$ in Eq.~\ref{eq:tau_q} and obtain,
\begin{align}
\tau_t = \frac{m^2}{\eta^n R'}.
\end{align}
Comparing $\tau_t$ with $\tau_c$ and following the same way for original boson-sampling in Sec.~\ref{sec:sampling_time}, we plot the critical photon-number to reach quantum supremacy as a function of $\eta$ for the time-bin encoding scheme.

@@

\subsection{Hybrid Encoding Architecture}

@@

From the perspective of the encoding scheme, the sampling speed (unitary transformation) of microwave a boson-sampler is much slower than that of spatially-encoding boson-samplers. As showed in Fig.~\ref{fig:curves_2}, to beat the same advanced classical computer, we actually need higher system efficiency or a larger number of photons if the microwave scheme is utilised. However, the advantage of the superconducting platform exactly lies in its marvelously high fidelities ($>0.90$) of generation and measurement processes \cite{bib:21}, to which no ordinary optical boson-sampler is comparable. Though with high processing speed, the effective repeat frequency of a spatially-encoded optical boson-sampler is greatly reduced by post-selection due to the limited collection and detection efficiency. \comment{This doesn't make sense to me - explain it better}

\subsection{Lossy Boson-Sampling}

In real-world experiments, photons are inevitably subject to loss and other errors, making it sample from an erroneous distribution rather than the desired boson-sampling distribution. As boson-sampling only comprises passive optical elements, it is not universal and there are no known fault tolerance or error-correction schemes. From the beginning, Aaronson and Arkhipov 
cite{bib:2} have already noted this problem. And in their original paper introducing boson-sampling, proved that boson-sampling is classically hard even if sampling from an approximate distribution with multiplicative error. Furthermore, several follow-up works have discussed the effect of realistic experimental imperfections in more detail, such as losses \cite{bib:22}, mode-mismatch due to calibration errors within the linear optical network \cite{bib:23, bib:24}, and photon distinguishability \cite{bib:25, bib:26, bib:27, bib:28, bib:29, bib:30}.

@@

It was also proved that calculating $\mathrm{Pr}(U_{S,T})$ is as least as hard as calculating the permanent of an \mbox{$m\times n$} submatrix of a Harr-random matrix. Let us introduce a modification of $\tau_q$ and $\tau_c$ in the context of boson-sampling with lost photons. The average sampling interval $\tau_q$ is obtained by replacing the probability of detecting $n$ photons in coincidence, $\eta^n$, in Eq.~\ref{eq:tau_q} with,
\begin{align}
P(n,n_\mathrm{los}) = \eta^n(1-\eta)^{n_\mathrm{lost}}\binom{n+n_
\mathrm{lost}}{n_\mathrm{lost}},
\end{align}
where $n$ is the number of photons after loss, and $n_\mathrm{lost}$ is the number of lost photons. So we have,
\begin{align}
\tau_q = \frac{1}{\eta^n(1-\eta)^{n_\mathrm{lost}}\binom{n+n_
\mathrm{lost}}{n_\mathrm{lost}}R}.
\end{align}

@@

\subsection{Scattershot Boson-Sampling}

The brightness of $n$ photons will decay exponentially as $p^n$ if each SPS has efficiency $p$. For current mature heralded SPSs based on spontaneous parametric down-conversion (SPDC), one solution is to introduce more SPDC sources and make $p$ approach unity by the technique of multiplexing \cite{bib:34}. The idea is simple -- we prepare a large bank of SPDC sources triggering in parallel, all feeding into a multiplexer, which routes successfully heralded photons to the desired outputs. This allows state preparation arbitrarily close to unity in principle (in the ideal lossless case), with a sufficient number of sources operating in parallel.

@@

\begin{figure}[!htb]
\includegraphics[width=\columnwidth]{scattershot_model}
\caption{The scheme for scattershot boson-sampling \cite{bib:36}. For an \mbox{$m\times m$}-mode interferometer, there are $m$ pairs of SPDCs, of which, upon post-selection, only $n$ generate single photons at a time.} \label{fig:scattershot_model}
\end{figure}

An intuitive explanation for the computational complexity of SBS is that in ordinary boson-sampling we are sampling from a distribution of permanents of Haar-random matrices. The SBS protocol simply adds an extra level of randomisation -- specifically, a random permutation -- prior to the Haar-random matrix. Intuitively, we obviously expect that a permutation of a random matrix yields another random matrix.

@@

As pointed out in Sec.~\ref{sec:sampling_time}, the only thing we care about in the competition of quantum boson-samplers versus classical simulators is who would be the first to produce a legitimate sample from the permanent-related probability distribution. And in the context of no a priori knowledge about which input configuration would be picked in the experiment, to simulate SBS, a classical computer can obtain a legitimate sample just by simulating ordinary boson-sampling with randomly permuted input states, which does not incur any computational overhead, i.e the `hard' part of the computation (calculating permanents) remains unchanged. Thus, we follow the same $t_c$ in Eq.~\ref{eq:tau_c} for SBS, which does not need to be multiplied by a combinatorial factor at all as in \cite{bib:37}.

>>>>

@@

\begin{abstract}
Boson-sampling is thus far the most feasible approach for first demonstrating the supremacy of quantum computation over classical computation in the near future. We take the view that proof-of-principle demonstration of quantum supremacy is a pressing goal, with enormous implications, and should be considered an utmost priority by the experimental community. Here we strive to outline a roadmap to achieving this elusive goal, examining competing architectures and technologies, and paving the way forward for the first experimental demonstration of post-classical computation. We firstly develop a practical model to compare the performance of boson-samplers and classical simulators. In this model, we find that low-photon-number boson-sampling operating in the current experimental regime (\mbox{$n<10$} photons) could beat present commodity computers with GHz clock-rates only if the overall system efficiency and repetition rate are sufficiently high. Further, we calculate the required efficiency and photon-number to surpass classical computers of different clock-rates using various experimental schemes. Especially, we find that only if both photon-number-resolving detectors and high-speed optical shutters are available, scattershot boson-sampling is feasibly scalable to tens of photons. In order to provide a clear roadmap for boson-sampling, we also review and discuss various state-of-the-art technologies for single-photon sources, quantum circuits, and single-photon detectors.
\end{abstract}

@@

Quantum computers are expected to be able to solve certain problems exponentially faster than classical computers \cite{bib:1}, and people have been making unremitting efforts towards them for more than two decades. However, actual quantum computers are presently still in their infancy. Thanks to its simplicity, boson-sampling is expected to be the first specialised quantum computer outperforming classical ones in the near term \cite{bib:2}. More importantly, this would also constitute the strongest evidence yet against a foundational tenet in computer science: the Extended Church-Turing Thesis (ECT), which postulates that all realistic physical systems can be efficiently simulated with a (classical) probabilistic Turing machine. However, the realisation of large-scale boson-samplers still faces many demanding technological challenges, despite being far easier to build than universal quantum computers.

To scale up sufficiently to exhibit quantum supremacy, the first and most important consideration is the total efficiency of the system, including the generation efficiency of single-photon sources, the coupling and propagation efficiency of the circuit, and the coupling and detection efficiency of single-photon detectors.

In Sec.~\ref{sec:sampling_time}, taking the repetition rate of the sampling system into account, we develop a rigorous and fair model for comparing the performance of boson-samplers and classical simulators. Given a certain classical computer, the regime of quantum supremacy of boson-sampling is specified by a simple continuous curve in the plane of total efficiencies versus photon-number. \comment{To our surprise, there is such a (optimal) turning point in this curve where four-photon boson-samplers with repetition rates of 76MHz could beat present commodity computers with GHz clock-rates}. Based upon this, such a computer could only be defeated by \mbox{$20\sim 30$}-photon boson-sampling \cite{bib:2, bib:3}. Furthermore, in Sec.~\ref{sec:exp_schemes}, we calculate the required efficiencies and photon-numbers for different experimental schemes to outperform classical computers of given clock-rates. In particular, we find that both photon-number-resolving detectors and high-speed optical shutters are indispensable to correctly implementing scattershot boson-sampling with tens of photons.

@@

\section{Sampling time} \label{sec:sampling_time}

@@

\subsection{Quantum simulation}

Boson-sampling is a process in which $n$ indistinguishable photons are injected into an $m$-mode linear optical interferometer, characterised by linear optics transformation $U$ (i.e beamsplitters and phase-shifters), and then detected at the output ports using photodetectors, as shown in Fig.~\ref{fig:model} \cite{bib:2}. $U$ evolves the photonic creation operators according to,
\begin{align}
\hat{U} \hat{a}^\dag_i \hat{U}^\dag \to \sum_{j=1}^m U_{i,j} \hat{a}^\dag_j,
\end{align}
where $\hat{a}^\dag_i$ is the photon creation operator for the $i$th mode.

Choosing the mode occupation-number configurations as the basis states, the input state can be written as a list of occupation numbers \mbox{$\ket{S}=\ket{s_1,\dots,s_m}$}, where $s_i$ is the number of photons in the $i$th input mode. As photon-number is preserved via linear optics transformations, the output state can be denoted in the same basis, but a superposition of all different output configurations,
\begin{align}
\ket\varphi = \sum_{i=1}^{|\Phi|}\alpha_i \ket{T_i},
\end{align}
where \mbox{$\ket{T_i} = \ket{t_1,\dots,t_m}$}, and with each basis state satisfying,
\begin{align}
\sum_{i=1}^m s_i = \sum_{i=1}^m t_i = n \,\,\forall\,\, S,T.
\end{align}
We define $\Phi$ as the set of all photon-number configuration basis vectors. It is not hard to prove that the total number of elements in $\Phi$ is,
\begin{align}
|\Phi|=\binom{m+n-1}{n}.
\end{align}
The amplitude associated with each output basis state is given by \cite{bib:2, bib:12},
\begin{align}
\bra{T_i} \Phi(U) \ket{S} = \frac{\mathrm{Per}(U_{S,T_i})}{\sqrt{s_1!\dots s_m!t_1!\dots t_m!}},
\end{align}
where $\Phi(U)$ is a homomorphism of $U$, and $U_{S,T_i}$ is an $n\times n$ submatrix of $U$, obtained by taking $s_i$ copies of the $i$th row of $U$, and $t_j$ copies of the $j$th column of $U$. It has been proven that \cite{bib:2, bib:13, bib:14} one will observe a collision (i.e. two or more photons in the same mode) with probability bounded away from zero only if the number of modes scales as at least $m=O(n^2)$ -- the so-called `bosonic birthday paradox' \cite{???}. In this instance, it suffices to use non-photon-number-resolving detectors, substantially simplifying technological requirements. Thus, throughout this paper we only discuss the `collision-free' situation, i.e. $s_i,t_j=\{0,1\}$.

@@

\subsection{Classical simulation}

How can a classical computer simulate this bosonic process, and sample from its exact or approximate probability distribution? A na{\"i}ve brute-force method is to calculate all $|\Phi|$ permanents of the set of submatrices $U_{S,T_i}$, thereby determining the full probability distribution, and then sample from this known distribution. \textbf{\comment{However, neither does the boson-sampler know even the approximate probability distribution before enough samples are obtained}} \comment{What does that sentence mean?}. Smarter classical algorithms exist to generate correct samples without knowing the entire underlying distribution. For example, using the acceptance-rejection method \cite{bib:15} \comment{Is there a better reference than Wikipedia?} one can sample an output configuration $\ket{T_i}$ uniformly from the configuration space and a number $u$ from the uniform distribution over $\{0,1\}$, then calculate the permanents of the respective submatrices and compare them to $u$; If \mbox{$u<|\mathrm{Per}(U_{S,T_i})|^2$}, we accept the sample $\ket{T_i}$; If not, it is rejected, and we repeat. Such methods could be successful with probability $O(1)$. One sample only requires calculating one permanent, significantly better than the brute-force method of calculating the entire distribution.

The permanent is a matrix function originating from the permutation symmetry of bosons, similar to the role of determinants for fermions. Conceptually, the permanent effectively sums over the amplitudes of all possible paths for photons in given input configurations to yield a given output configuration. Na{\"i}ve calculation of the permanent of an $n\times n$ matrix requires $O(n!)$ operations. However, a more efficient classical algorithm developed by Ryser \cite{bib:16} in 1963 requires $O(n^2 2^{n+1})$ operations, which is unfortunately nonetheless exponential. No general polynomial-time algorithms have been described, in contrast to the clever Gaussian elimination method for calculating matrix determinants. In fact, calculating the permanent was proven to be \#P-complete in 1979 by Valiant \cite{bib:17}.

@@

Comparing the requisite times, we plot the critical photon-number as a function of efficiency, $\eta$, for the quantum boson-sampler to beat its classical counterpart, as shown in Fig.~\ref{fig:curves_1}(a). The right part of each line is the area of quantum supremacy. For a given single-photon efficiency, with increasing $n$, the quantum simulator's superiority becomes more prominent. Thus, scaling up the size of boson-sampling is essential to achieving this goal. But there is an upper limit. Taking \mbox{$\eta=0.4$} as an example, it is impossible to reach quantum supremacy for a classical computer of petaFLOPS by only increasing the number of single photons. A more convenient approach is to increase the efficiency of the quantum system. As the efficiency increases, the minimum photon-number required to achieve quantum supremacy decreases, simplifying experimental resource requirements.

@@

The repetition rate, $R$, is another important factor affecting the performance of quantum simulators that we must consider in practical experiments. If $\eta^n$ is taken as the post-selection success probability of a single trial, then \mbox{$\eta^n R$} indicates how many successful runs we can implement per second. For example, to beat a gigaFLOPS classical computer, 4-photon boson-sampling with \mbox{$R=76\mathrm{MHz}$} is sufficient if the total efficiency is \mbox{$\eta=0.4$}. As shown in Fig.~\ref{fig:curves_1}(b), however, \mbox{$\eta=0.2$} and \mbox{$n=2$} is sufficient if the repetition rate increases to 1GHz. In other words, by increasing the repetition rate, the efficiency and photon-number requirements are relaxed. However, we cannot increase the repetition rate indefinitely. It is restricted by the pulse rate of the SPS, experimental architecture, and dead-time of the detectors. In the next section we discuss the impact of the choice of experimental architecture.

Here we must point out two things. First, Ryser's algorithm is actually not faster than using the original formula to calculate the matrix permanent when photon number is small (\mbox{$n<6\sim 7$}). Thus, the above conditions may not be tight. For example, the efficiency must be larger than $\sim 0.6$ if two- (four)-photon boson-sampling with a repetition rate of 1GHz (76MHz) is expected to beat a gigaFLOPS classical computer.

Why then should we nonetheless scale up to more photons (\mbox{$n=20\sim 30$}), even though boson-sampling with relatively few photons could beat current prevailing personal computers in realistic experimental regimes?

One reason is that computer scientists would very much like to verify the scaling in the complexity of such devices. Are they really exponentially faster? While no finite experiment can make a definitive conclusion about the scaling of computational resources -- since these are asymptotic statements -- if very large-scale devices were realisable, and verification techniques reliable, one could make successively stronger statements about asymptotic scaling as the size of the system increased. Of particular interest to computer scientists is the validity of the ECT thesis. While no experiment of finite size can confirm or dispute the ECT thesis, verifiable large-scale boson-sampling could provide circumstantial evidence for the existence of physical systems that cannot be efficiently classically simulated.

Another reason is that people have not physically constructed relatively large multi-photon interferometers yet. The theory of boson-sampling still faces the risk of failure and quantum devices may break down with more photons.

Finally, although no practical uses for boson-sampling have been described, it is entirely plausible that useful applications may be found in the future. Were that the case, scaling up boson-sampling experiments would similarly scale up the size of the problems it could solve efficiently.

\section{Experimental schemes} \label{sec:exp_schemes}

\subsection{Spatial encoding architecture}

A fully spatially-encoded scheme is one whereby the circuit itself does not consume any time resources (beyond the propagation time from the input to the output of the circuit, which is fixed). This includes optical waveguides, fused fibre couplers and bulk-optics circuits, as shown in Fig.~\ref{fig:space_encoding}. The time required for this scheme is the same as Eq.~\ref{eq:tau_q}.

@@

\subsection{Time-bin encoding architecture}

@@

\begin{figure}[!htb]
\includegraphics[width=0.7\columnwidth]{loop_arch}
\caption{The fibre-loop scheme for boson-sampling. A pulse-train of photons in $m$ time-bins defines the input state, with time-bin separation $\tau$. The pulse-train is coupled into and out of the outer loop using two outer dynamic switches. The inner loop of length $\tau$ allows neighbouring time-bins to interfere at the central dynamic switch. The outer two switches need only switch between being completely reflective and completely transmissive, whereas the inner switch must be able to tune to any beamsplitter reflectivity. The only component that scales with the size of the interferometer is the length of the outer loop, which must be long enough to contain the entire pulse-train, $>m\tau$.} \label{fig:loop_arch}
\end{figure}

In this design, a sequence of single photons is injected and interfere with one other through the time delay and fibre beamsplitter. As in Fig.~\ref{fig:loop_arch}, a small fibre loop of round-trip time \mbox{$\tau=1/R$} acts as the time delay, where the node is a dynamic beam splitter for interference. The small loop is embedded in a large fibre loop of travelling time $>m\tau$, which is used to return the photon pulse-train back for the next round of interference. These two embedded fibre loops together fully implement the interference network, encoding entirely in the temporal degree of freedom. According to Reck's result, one needs $O(m^2)$ beamsplitters to construct an arbitrary \mbox{$m\times m$} interferometer. With $m$ time-bins, the beamsplitter works $m$ times in each round trip. So $m$ round trips are required to implement an arbitrary unitary transformation. After $m$ loops of evolution, the photons are all coupled out of the loop for time-resolved coincidence measurements.

As described above, the circuit itself consumes time resources $m^2\tau$, thus the effective repetition rate is,
\begin{align}
\frac{1}{m^2\tau}=\frac{R'}{m^2},
\end{align}
where $R'$ is the repetition frequency of the photonic time-bins, which we use instead of $R$ in Eq.~\ref{eq:tau_q} to obtain,
\begin{align}
\tau_t = \frac{m^2}{\eta^n R'}.
\end{align}
Comparing $\tau_t$ with $\tau_c$ and \comment{following the same way for original boson-sampling in Sec.~\ref{sec:sampling_time} \textbf{What does this mean?}}, we plot the critical photon-number to reach quantum supremacy as a function of $\eta$ for the time-bin encoding scheme.

@@

\subsection{Hybrid encoding architecture}

@@

From the perspective of the encoding scheme, the sampling speed (unitary transformation) of the microwave boson-sampler is much slower than that of spatially-encoding boson-samplers. As shown in Fig.~\ref{fig:curves_2}, to beat the same advanced classical computer, we actually need higher system efficiency or a larger number of photons if the microwave scheme is utilised. However, the advantage of the superconducting platform exactly lies in its marvelously high fidelities ($>0.90$) of generation and measurement processes \cite{bib:21}, to which no ordinary optical boson-sampler is comparable. Though with high processing speed, the effective repeat frequency of a spatially-encoded optical boson-sampler is greatly reduced by post-selection due to the limited collection and detection efficiency. \comment{This doesn't make sense to me - explain it better}

\subsection{Lossy boson-sampling}

In real-world experiments, photons are inevitably subject to loss and other errors, making it sample from an erroneous distribution rather than the desired boson-sampling distribution. As boson-sampling only comprises passive optical elements, it is not universal and there are no known fault tolerance or error-correction schemes. From the beginning, Aaronson and Arkhipov 
cite{bib:2} noted this problem. And in their original paper introducing boson-sampling, proved that boson-sampling is classically hard even if sampling from an approximate distribution with multiplicative error. Furthermore, several follow-up works have discussed the effect of realistic experimental imperfections in more detail, such as losses \cite{bib:22}, mode-mismatch due to calibration errors within the linear optical network \cite{bib:23, bib:24}, and photon distinguishability \cite{bib:25, bib:26, bib:27, bib:28, bib:29, bib:30}.

@@

It was also proven that calculating $\mathrm{Pr}(U_{S,T})$ is as least as hard as calculating the permanent of an \mbox{$m\times n$} submatrix of a Haar-random matrix. Let us introduce a modification of $\tau_q$ and $\tau_c$ in the context of boson-sampling with lost photons. The average sampling interval $\tau_q$ is obtained by replacing the probability of detecting $n$ photons in coincidence, $\eta^n$, in Eq.~\ref{eq:tau_q} with,
\begin{align}
P(n,n_\mathrm{lost}) = \eta^n(1-\eta)^{n_\mathrm{lost}}\binom{n+n_\mathrm{lost}}{n_\mathrm{lost}},
\end{align}
where $n$ is the number of photons after loss, and $n_\mathrm{lost}$ is the number of lost photons. So we have,
\begin{align}
\tau_{q(\mathrm{lost})} = \frac{1}{\eta^n(1-\eta)^{n_\mathrm{lost}}\binom{n+n_
\mathrm{lost}}{n_\mathrm{lost}}R}.
\end{align}

@@

\subsection{Scattershot boson-sampling}

The brightness of $n$ photons decays exponentially as $p^n$ if each SPS has efficiency $p$. For current mature heralded SPSs based on spontaneous parametric down-conversion (SPDC), one solution is to introduce more SPDC sources and make $p$ approach unity using multiplexing \cite{bib:34}. The idea is simple -- we prepare a large bank of SPDC sources triggering in parallel, all feeding into a multiplexer, which routes successfully heralded photons to the desired outputs. This allows state preparation arbitrarily close to unity in principle (in the ideal lossless case), with a sufficient number of sources operating in parallel.

@@

\begin{figure}[!htb]
\includegraphics[width=\columnwidth]{scattershot_model}
\caption{The scheme for scattershot boson-sampling \cite{bib:36}. For an \mbox{$m\times m$}-mode interferometer, there are $m$ SPDCs non-deterministically preparing photon-pairs, of which, upon post-selection, only $n$ generate single photons at a time.} \label{fig:scattershot_model}
\end{figure}

An intuitive explanation for the computational complexity of SBS is that in ordinary boson-sampling we are sampling from a distribution of permanents of Haar-random matrices. The SBS protocol simply adds an extra level of randomisation -- specifically, a random permutation -- prior to the Haar-random matrix. Intuitively, we obviously expect that a permutation of a random matrix yields another random matrix. Thus, all the same complexity proofs hold.

@@

As pointed out in Sec.~\ref{sec:sampling_time}, the only thing we care about in the competition between quantum boson-samplers and classical simulators is who would be the first to produce a legitimate sample from the permanent-related probability distribution. And in the context of no a priori knowledge about which input configuration would be picked in the experiment, to simulate SBS, a classical computer can obtain a legitimate sample just by simulating ordinary boson-sampling with randomly permuted input states, which does not incur any computational overhead, i.e the `hard' part of the computation (calculating permanents) remains unchanged. Thus, we can utilise the same $t_c$ as in Eq.~\ref{eq:tau_c} for SBS, which does not need to be multiplied by a combinatorial factor as in \cite{bib:37}.
